Generating the Blueprints of the Java Ecosystem

### 2015-02-28 msr.tex
6 and 7

\begin{figure}
\centering
\includegraphics[scale=0.45]{import-process}
\caption{The dataset construction process}
\label{fig:dataset-construction}
\end{figure}

@@

\begin{table}
\centering
\caption{The selected Maven projects' size metrics}
\label{tbl:oss-size-metrics}
\begin{tabular}{l r}
 \hline
\textbf{Metric} & \textbf{Value}\\
\hline
Project Count & 11,365\\
File Count & 449,213\\
Package Count & 59,436\\
Lines of Code & 74,565,772\\
Source Lines of Code & 40,921,287\\
Comment Lines of Code & 25,268,959\\
Number of Classes & 370,518\\
Number of Interfaces & 66,352\\
Number of Enumerations & 9,879\\
Measurement Count & 32,844,836\\
\hline
\end{tabular}
\end{table}

@@

\begin{table}
\centering
\caption{List of selected DSL application libraries}
\label{tbl:dsl-list}
\begin{tabular}{l l}
 \hline
\textbf{DSL} & \textbf{Java Package}\\
\hline
Regular Expressions & \verb|java.util.regex|\\
XML & \verb|javax.xml|, \verb|org.w3c| and \verb|org.xml|\\
SQL & \verb|java.sql| and \verb|javax.sql|\\
XPath & \verb|java.xml.xpath|\\
XSLT & \verb|javax.xml.transform|\\
RTF & \verb|javax.swing.text.rtf|\\
HTML & \verb|javax.swing.text.html|\\
\hline
\end{tabular}
\end{table}

@@

\begin{table}
\centering
\caption{Class Design Metrics}
\label{tbl:class-selected-metrics}
\begin{tabular}{l l}
 \hline
\multicolumn{2}{l}{\textit{\textbf{Class Design}}}\\
\hline
Depth Of Inheritance Tree & \textit{ckjm}\\
Coupling Between Objects & \textit{ckjm}\\
Weighted Methods Per Class & \textit{ckjm}\\
Response For Class & \textit{ckjm}\\
Lack Of Cohesion In Methods & \textit{ckjm}\\
Number Of Children & \textit{ckjm}\\
Attribute Hiding Factor & \textit{clmt}\\
Coupling Between Methods & \textit{ckjm}\\
Average Method Complexity & \textit{ckjm}\\
Cohesion Among Methods of Class & \textit{ckjm}\\
Data Access Metric & \textit{ckjm, clmt}\\
Inheritance Coupling & \textit{ckjm}\\
Lack Of Cohesion In Methods3 & \textit{ckjm}\\
Measure Of Aggregation & \textit{ckjm}\\
Measure Of Functional Abstraction & \textit{ckjm}\\
Number of Attributes & \textit{clmt}\\
Method Hiding Factor & \textit{clmt}\\
\hline
\end{tabular}
\end{table}

\begin{table}
\centering
\caption{Method Design Metrics}
\label{tbl:method-selected-metrics}
\begin{tabular}{l l}
 \hline
\multicolumn{2}{l}{\textit{\textbf{Method Design}}}\\
\hline
Number of Method Parameters & \textit{clmt}\\
Number of Methods & \textit{clmt, ckjm}\\
McCabe Cyclomatic Complexity & \textit{ckjm, clmt}\\
\hline
\end{tabular}
\end{table}

@@

\begin{figure*}
\centering
\includegraphics[scale=0.7]{database-schema}
\caption{The database schema}
\label{fig:database-schema}
\end{figure*}

@@

\begin{table}
\centering
\caption{DSL Usage Combinations Ranking}
\label{tbl:dsl-top-usage}
\begin{tabular}{l r}
 \hline
\textbf{DSLs} & \textbf{Count}\\
\hline
XML & 1,561\\
Regex & 909\\
SQL & 493\\
XML, XSLT & 475\\
Regex, XML, XSLT & 158\\
Regex, XML & 303\\
SQL, XML & 162\\
Regex, SQL & 116\\
Regex, SQL, XML, XSLT & 80\\
Regex, SQL, XML & 71\\
SQL, XML, XSLT & 54\\
XML, XPath & 50\\
\hline
\end{tabular}
\end{table}

The initial goal of this experiment was to provide quantifiable results that are indicative regarding the usage of {\sc dsl}s in multiple Java projects. The related metrics can be found in the database table named \textit{measuremnt\_type}, and they are the following: \textit{RTFUsage\_clmt}, \textit{RegexUsage\_clmt},
\textit{HTMLUsage\_clmt}, \textit{HTMLUsage\_clmt}, \textit{XPathUsage\_clmt}, \textit{XSLTUsage\_clmt}, \textit{XMLUsage\_clmt}, \textit{SQLUsage\_clmt}, and \textit{DSLCount\_clmt}. With a python script that analyses the measurements stored in the database, we produced Table~\ref{tbl:dsl-top-usage}. This table, lists popular {\sc dsl} usage combinations found in our dataset. An interesting observation is that {\sc xml} is the most widespread {\sc dsl} with 1561 occurrences in 11365 projects (13\%). Regular expressions are also popular with 909 occurrences (7\%). Note that the numbers represents the projects where {\sc xml} and regular expressions were used as the only {\sc dsl}.

@@

This dataset adopted a narrow selection process, identifying and analysing only projects that were written in Java and had both the source code and the binary {\sc jar} available in the repository. The latter reduced significantly the number of projects that were analysed, since many of them provided only the binary {\sc jar}. Since only \textit{clmt} analyses the source code, this ended in drastically reduction of measurements that were exposing object-oriented and package design issues.

In addition, only one version was analysed per project. Usually, the latest, which was replaced by a prior version, only if it was violating the aforementioned selection process. This decision renders the dataset unusable, for research focused on software evolution.

@@

In reference \cite{MKLGS14}, Mitropoulos et al. also experimented in a similar way, by running the FindBugs~\cite{HP04} tool on a large part of the maven repository. Our work differs from these two approaches since it presents the metrics calculated by the three different static analysis tools.

>>>>

\begin{figure}
\centering
\includegraphics[scale=0.45]{import-process}
\caption{The dataset construction process.}
\label{fig:dataset-construction}
\end{figure}

@@

\begin{table}
\centering
\caption{The selected Maven projects' size metrics.}
\label{tbl:oss-size-metrics}
\begin{tabular}{l r}
 \hline
\textbf{Metric} & \textbf{Value}\\
\hline
Project Count & 11,365\\
File Count & 449,213\\
Package Count & 59,436\\
Lines of Code & 74,565,772\\
Source Lines of Code & 40,921,287\\
Comment Lines of Code & 25,268,959\\
Number of Classes & 370,518\\
Number of Interfaces & 66,352\\
Number of Enumerations & 9,879\\
Measurement Count & 32,844,836\\
\hline
\end{tabular}
\end{table}

@@

\begin{table}
\centering
\caption{List of selected {\sc dsl} application libraries.}
\label{tbl:dsl-list}
\begin{tabular}{l l}
 \hline
\textbf{DSL} & \textbf{Java Package}\\
\hline
Regular Expressions & \verb|java.util.regex|\\
XML & \verb|javax.xml|, \verb|org.w3c| and \verb|org.xml|\\
SQL & \verb|java.sql| and \verb|javax.sql|\\
XPath & \verb|java.xml.xpath|\\
XSLT & \verb|javax.xml.transform|\\
RTF & \verb|javax.swing.text.rtf|\\
HTML & \verb|javax.swing.text.html|\\
\hline
\end{tabular}
\end{table}

@@

\begin{table}
\centering
\caption{Class Design Metrics.}
\label{tbl:class-selected-metrics}
\begin{tabular}{l l}
 \hline
\multicolumn{2}{l}{\textit{\textbf{Class Design}}}\\
\hline
Depth Of Inheritance Tree & \textit{ckjm}\\
Coupling Between Objects & \textit{ckjm}\\
Weighted Methods Per Class & \textit{ckjm}\\
Response For Class & \textit{ckjm}\\
Lack Of Cohesion In Methods & \textit{ckjm}\\
Number Of Children & \textit{ckjm}\\
Attribute Hiding Factor & \textit{clmt}\\
Coupling Between Methods & \textit{ckjm}\\
Average Method Complexity & \textit{ckjm}\\
Cohesion Among Methods of Class & \textit{ckjm}\\
Data Access Metric & \textit{ckjm, clmt}\\
Inheritance Coupling & \textit{ckjm}\\
Lack Of Cohesion In Methods3 & \textit{ckjm}\\
Measure Of Aggregation & \textit{ckjm}\\
Measure Of Functional Abstraction & \textit{ckjm}\\
Number of Attributes & \textit{clmt}\\
Method Hiding Factor & \textit{clmt}\\
\hline
\end{tabular}
\end{table}

\begin{table}
\centering
\caption{Method Design Metrics.}
\label{tbl:method-selected-metrics}
\begin{tabular}{l l}
 \hline
\multicolumn{2}{l}{\textit{\textbf{Method Design}}}\\
\hline
Number of Method Parameters & \textit{clmt}\\
Number of Methods & \textit{clmt, ckjm}\\
McCabe Cyclomatic Complexity & \textit{ckjm, clmt}\\
\hline
\end{tabular}
\end{table}

@@

\begin{figure*}
\centering
\includegraphics[scale=0.65]{database-schema}
\caption{The database schema.}
\label{fig:database-schema}
\end{figure*}

@@

\begin{table}
\centering
\caption{{\sc dsl} Usage Combinations Ranking.}
\label{tbl:dsl-top-usage}
\begin{tabular}{l r}
 \hline
\textbf{DSLs} & \textbf{Count}\\
\hline
XML & 1,561\\
Regex & 909\\
SQL & 493\\
XML, XSLT & 475\\
Regex, XML, XSLT & 158\\
Regex, XML & 303\\
SQL, XML & 162\\
Regex, SQL & 116\\
Regex, SQL, XML, XSLT & 80\\
Regex, SQL, XML & 71\\
SQL, XML, XSLT & 54\\
XML, XPath & 50\\
\hline
\end{tabular}
\end{table}

The initial goal of this experiment was to provide quantifiable results that are indicative regarding the usage of {\sc dsl}s in multiple Java projects. The related metrics can be found in the database table named \textit{measuremnt\_type}, and they are the following: \textit{RTFUsage\_clmt}, \textit{RegexUsage\_clmt},
\textit{HTMLUsage\_clmt}, \textit{HTMLUsage\_clmt}, \textit{XPathUsage\_clmt}, \textit{XSLTUsage\_clmt}, \textit{XMLUsage\_clmt}, \textit{SQLUsage\_clmt}, and \textit{DSLCount\_clmt}. With a python script that analyses the measurements stored in the database, we produced the results illustrated in Table~\ref{tbl:dsl-top-usage}. This table, lists popular {\sc dsl} usage combinations found in our dataset. An interesting observation is that {\sc xml} is the most widespread {\sc dsl} with 1561 occurrences in 11365 projects (13\%). Regular expressions are also popular with 909 occurrences (7\%). Note that the numbers represents the projects where {\sc xml} and regular expressions were used as the only {\sc dsl}.

@@

As we mentioned earlier, to produce our dataset we adopted a narrow selection process. In particular, we identified and analyzed only projects that were written in Java and had both the source code and the binary {\sc jar} available in the repository. The latter reduced significantly the number of projects that were analyzed, because many of them provided only the binary {\sc jar}. Since only \textit{clmt} analyses the source code, this resulted in the reduction of the number of measurements that exposed object-oriented and package design issues.

Finally, only one version was analyzed per project (the latest, unless it violated the aforementioned restriction).
This decision renders the dataset unusable, for research focused on software evolution.

@@

Also, Mitropoulos et al. performed a similar experiment~\cite{MKLGS14}. In particular, they used the FindBugs~\cite{HP04} static analysis tool, to analyze a large part of the maven repository. Their dataset includes for each {\sc jar}, a corresponding bug collection produced by FindBugs. Our work differs from these two approaches since it presents the metrics calculated by the three different static analysis tools that calculate metrics regarding object-oriented design, program size, and package design.
